---
title: "Diffusion Models of Pragmatic Inference Tasks"
author: "Dan Yurovsky & Erica Yoon"
date: "January 07, 2014"
output: html_document
---
`r library(knitr)`
`r opts_chunk$set(message=FALSE, warning=FALSE)`

## Data Loading

Load required Libraries
```{r}
rm(list=ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(magrittr)
library(directlabels)
library(RWiener)
```

Fixing up the wiener_plot function to deal with missing data
```{r}
wiener.plot <- function (dat) 
{
    rt = as.double(dat$q)
    rc = as.numeric(as.factor(dat$resp))
    dpos = tryCatch(density(rt[rc == 1], from = 0),error=function(e) NA)
    dneg = tryCatch(density(rt[rc == 2], from = 0),error=function(e) NA)
    maxt = max(pretty(max(rt)))
    
    maxd <- NA
    if(is.na(dpos[1])){
      maxd <- max(dneg$y)
      } else if(is.na(dneg[1])){
        maxd <- max(dpos$y)
        } else {
          maxd <- max(dpos$y, dneg$y)
          }
    
    par(mar = c(0, 5, 0, 0), mfcol = c(2, 1), ask = FALSE)
    plot(dpos, xlim = c(0, maxt), ylim = c(0, maxd), las = 2, 
        lwd = 2, col = "green3", main = "", ylab = "", ask = FALSE)
    rug(rt[rc == 1], col = "green3")
    mtext("Density of positive responses", side = 2, line = 4, 
        cex = 0.8)
    plot(dneg, xlim = c(0, maxt), ylim = c(maxd, 0), las = 2, 
        lwd = 2, col = "red", main = "", ylab = "", ask = FALSE)
    mtext("Density of negative responses", side = 2, line = 4, 
        cex = 0.8)
    rug(rt[rc == 2], col = "red", side = 3)
}
```

Read in adult data
```{r,fig.height=3,fig.width=4.5}
data <- read.csv('150105-simpimpsc-turk_complete.csv') %>%
  filter(trialtype %in% c("control_double", "inference")) %>%
  group_by(WorkerId,trialnum) %>%
  mutate(side = if((trialtype == "control_double") & (correct == "Y")) "more"
         else if((trialtype == "control_double") & (correct == "N")) "less"
         else if((trialtype == "inference") & (correct == "N")) "more"
         else "less") %>%
  mutate(resp = as.character(factor(side,levels=c("more","less"),
                                    labels=c("upper","lower")))) %>%
  mutate(q = rt/1000) %>%
  group_by() %>%
  select(q,trialtype,itemnum,resp) %>%
  # filter rt, remove rt outside of stdev*2
  filter(log(q) < mean(log(q)) + 2 * sd(log(q)),
         log(q) > mean(log(q)) - 2 * sd(log(q))) 

control.2v1.data <- filter(data,trialtype == "control_double",
                           itemnum=="2vs1") %>%
  select(q,resp) %>%
  as.data.frame

wiener.plot(control.2v1.data)

control.3v1.data <- filter(data,trialtype == "control_double",
                           itemnum=="3vs1") %>%
  select(q,resp) %>%
  as.data.frame

wiener.plot(control.3v1.data)

inference.2v1.data <- filter(data,trialtype == "inference",
                             itemnum=="2vs1") %>%
  select(q,resp) %>%
  as.data.frame

wiener.plot(inference.2v1.data)

inference.3v1.data <- filter(data,trialtype == "inference",
                             itemnum=="3vs1") %>%
  select(q,resp) %>%
  as.data.frame

wiener.plot(inference.3v1.data)
```

Compute parameters for each condition separately
```{r}
optim.control.2v1 <- optim(c(1, .1, .1, 1), wiener_deviance, 
                           dat=control.2v1.data, method="Nelder-Mead")

optim.control.3v1 <- optim(c(1, .1, .1, 1), wiener_deviance, 
                           dat=control.3v1.data, method="Nelder-Mead")

optim.inference.2v1 <- optim(c(1, .1, .1, 1), wiener_deviance, 
                           dat=inference.2v1.data, method="Nelder-Mead")

optim.inference.3v1 <- optim(c(1, .1, .1, 1), wiener_deviance, 
                           dat=inference.3v1.data, method="Nelder-Mead")

indiv.pars <- as.data.frame(rbind(optim.control.2v1$par,
                                  optim.control.3v1$par,
                                  optim.inference.2v1$par,
                                  optim.inference.3v1$par))
names(indiv.pars) <- c("Separation","Non.Decision","Bias","Drift")
indiv.pars$Condition <- c("Control", "Control", "Inference", "Inference")
indiv.pars$Item.Num<- c(2, 3, 2, 3)
print(indiv.pars)
```

Pin parameters for item.num together
```{r}
# Function for tying parameters together across conditions
many_drifts <- function(x, datlist) {
  l = 0
  for (c in 1:length(datlist)) {
    l = l + wiener_deviance(x[c(1, 2, 3, c+3)], datlist[[c]])
  }
  
  return(l)
}

datlist.2v1 <- list(control.2v1.data, inference.2v1.data)
datlist.3v1 <- list(control.3v1.data, inference.3v1.data)
# use nlm to estimate parameters
optim.2v1 <- optim(p=c(1, .1, .1, 1, 1), many_drifts, 
                   dat=datlist.2v1,method="Nelder-Mead")
optim.3v1 <- optim(p=c(1, .1, .1, 1, 1), many_drifts, 
                   dat=datlist.3v1,method="Nelder-Mead")

joint.pars <- as.data.frame(rbind(optim.2v1$par,optim.3v1$par))
names(joint.pars) <- c("Separation","Non.Decision","Bias",
                       "Control", "Inference")
joint.pars <- gather(joint.pars,Condition,Drift,Control:Inference)
joint.pars$Item.Num <- c(2,3,2,3)
joint.pars %<>% select(Separation,Non.Decision,Bias,
                       Drift,Condition,Item.Num) %>%
  arrange(Condition,Item.Num)

print(joint.pars)
```

Pin non-decision time for a group, pin boundaries and bias within a condition
```{r}
# Function for tying parameters together across conditions
multi.loss.func <- function(x, datmat) {
  l = 0
  num.conds <- length(datmat)
  num.types <- length(datmat[[1]])
  
  # model takes (sep, non-decision, bias, drift)
  
  # x has (non-decision, sep-1, bias-1, drift-11, drift-12,
  #                      sep-2, bias-2, drift-21, drift-22)

  for (cond in 1:num.conds) {
    for(type in 1:num.types){

      l = l + wiener_deviance(x[c(2+(cond-1)*(2+num.types), 1, 
                                  3+(cond-1)*(2+num.types), 
                                  3+(cond-1)*(2+num.types)+type)], 
                              datmat[[cond]][[type]])
    }
  }
  
  return(l)
}

#Function for reconstructing the parameter matrix
make.par.mat <- function(optim.output) {
  pars <- as.data.frame(t(optim.output$par))
  names(pars) <- c("NonDecision","Separation.2v1","Bias.2v1",
                   "Drift.Control.2v1","Drift.Inference.2v1",
                   "Separation.3v1","Bias.3v1",
                   "Drift.Control.3v1","Drift.Inference.3v1")

  par.mat <- expand.grid(Condition = c("2v1", "3v1"),
                         Trial.Type = c("Control", "Inference"))
  
  par.mat$NonDecision <- pars[,"NonDecision"]
  par.mat$Separation <- rep(c(pars[,"Separation.2v1"],
                              pars[,"Separation.3v1"]))
  par.mat$Bias <- rep(c(pars[,"Bias.2v1"],
                        pars[,"Bias.3v1"]))
  par.mat$Drift <- c(pars[,"Drift.Control.2v1"],
                     pars[,"Drift.Control.3v1"],
                     pars[,"Drift.Inference.2v1"],
                     pars[,"Drift.Inference.3v1"])
  
  return(par.mat)
}
```

```{r}
datlist.2v1 <- list(control.2v1.data, inference.2v1.data)
datlist.3v1 <- list(control.3v1.data, inference.3v1.data)
datmat <- list(datlist.2v1,datlist.3v1)

  # x has (non-decision, sep-1, bias-1, drift-11, drift-12,
  #                      sep-2, bias-2, drift-21, drift-22)

optim.adult <- optim(p=c(.1, 1,.5, 1, -1, 1, .5, 1, -1), multi.loss.func, 
                   dat=datmat,method="BFGS")

adult.pars <- make.par.mat(optim.adult)

print(adult.pars)
```

```{r,fig.height=3,fig.width=4.5}
kid.data <- read.csv('simpimpSCresults_141209_1.csv') %>%
  filter(trialtype %in% c("control_double", "inference")) %>%
  group_by(subid,trialnum) %>%
  mutate(side = if((trialtype == "control_double") & (correct == "Y")) "more"
         else if((trialtype == "control_double") & (correct == "N")) "less"
         else if((trialtype == "inference") & (correct == "N")) "more"
         else "less") %>%
  mutate(resp = as.character(factor(side,levels=c("more","less"),
                                    labels=c("upper","lower")))) %>%
  mutate(q = reaction.rt/1000) %>%
  group_by() %>%
  select(q,trialtype,itemnum,resp) %>%
  # filter rt, remove rt outside of stdev*2
  filter(log(q) < mean(log(q)) + 2 * sd(log(q)),
         log(q) > mean(log(q)) - 2 * sd(log(q)))

kid.control.2v1.data <- filter(kid.data,trialtype == "control_double",
                               itemnum=="2vs1") %>%
  select(q,resp) %>%
  as.data.frame

wiener.plot(kid.control.2v1.data)

kid.control.3v1.data <- filter(kid.data,trialtype == "control_double",
                               itemnum=="3vs1") %>%
  select(q,resp) %>%
  as.data.frame

wiener.plot(kid.control.3v1.data)

kid.inference.2v1.data <- filter(kid.data,trialtype == "inference",
                                 itemnum=="2vs1") %>%
  select(q,resp) %>%
  as.data.frame

wiener.plot(kid.inference.2v1.data)

kid.inference.3v1.data <- filter(kid.data,trialtype == "inference",
                                 itemnum=="3vs1") %>%
  select(q,resp) %>%
  as.data.frame

wiener.plot(kid.inference.3v1.data)
```

Compute kid parameters for each condition separately
```{r}
kid.optim.control.2v1 <- optim(c(1, .1, .1, 1), wiener_deviance, 
                           dat=kid.control.2v1.data, method="Nelder-Mead")

kid.optim.control.3v1 <- optim(c(1, .1, .1, 1), wiener_deviance, 
                           dat=kid.control.3v1.data, method="Nelder-Mead")

kid.optim.inference.2v1 <- optim(c(1, .1, .1, 1), wiener_deviance, 
                           dat=kid.inference.2v1.data, method="Nelder-Mead")

kid.optim.inference.3v1 <- optim(c(1, .1, .1, 1), wiener_deviance, 
                           dat=kid.inference.3v1.data, method="Nelder-Mead")

kid.indiv.pars <- as.data.frame(rbind(kid.optim.control.2v1$par,
                                      kid.optim.control.3v1$par,
                                      kid.optim.inference.2v1$par,
                                      kid.optim.inference.3v1$par))
names(kid.indiv.pars) <- c("Separation","Non.Decision","Bias","Drift")
kid.indiv.pars$Condition <- c("Control", "Control", "Inference", "Inference")
kid.indiv.pars$Item.Num<- c(2, 3, 2, 3)

print(indiv.pars)

print(kid.indiv.pars)
```

Pin parameters for item.num together
```{r}
# Function for tying parameters together across conditions
kid.datlist.2v1 <- list(kid.control.2v1.data, kid.inference.2v1.data)
kid.datlist.3v1 <- list(kid.control.3v1.data, kid.inference.3v1.data)
# use nlm to estimate parameters
kid.optim.2v1 <- optim(p=c(1, .1, .1, 1, 1), many_drifts, 
                   dat=kid.datlist.2v1,method="Nelder-Mead")
kid.optim.3v1 <- optim(p=c(1, .1, .1, 1, 1), many_drifts, 
                   dat=kid.datlist.3v1,method="Nelder-Mead")

kid.joint.pars <- as.data.frame(rbind(kid.optim.2v1$par,kid.optim.3v1$par))
names(kid.joint.pars) <- c("Separation","Non.Decision","Bias",
                       "Control", "Inference")
kid.joint.pars <- gather(kid.joint.pars,Condition,Drift,Control:Inference)
kid.joint.pars$Item.Num <- c(2,3,2,3)
kid.joint.pars %<>% select(Separation,Non.Decision,Bias,Drift,
                           Condition,Item.Num) %>%
  arrange(Condition,Item.Num)

print(kid.joint.pars)
```

```{r}
kid.datlist.2v1 <- list(kid.control.2v1.data, kid.inference.2v1.data)
kid.datlist.3v1 <- list(kid.control.3v1.data, kid.inference.3v1.data)
kid.datmat <- list(kid.datlist.2v1,kid.datlist.3v1)

  # x has (non-decision, sep-1, bias-1, drift-11, drift-12,
  #                      sep-2, bias-2, drift-21, drift-22)

optim.kid <- optim(p=c(.1, 1,.5, 1, -1, 1, .5, 1, -1), multi.loss.func, 
                   dat=kid.datmat,method="BFGS")

kid.pars <- make.par.mat(optim.kid)

print(kid.pars)

print(adult.pars)
```

Some copy/pasted code from the paper
```{r}
# 
# set.seed(0)
# dat <- rwiener(n=100, alpha=2, tau=.3, beta=.5, delta=.5)
# 
# dat2 <- rwiener(n=100, alpha=2, tau=.3, beta=.5, delta=.75)
# 
# dwiener(dat$q[1], alpha=2, tau=.3, beta=.5, delta=.5, resp=dat$resp[1], give_log=FALSE)
# 
# curve(dwiener(x, 2, .3, .5, .5, rep("upper", length(x))),
# xlim=c(0,3), main="Density of upper responses",
# ylab="density", xlab="quantile")
# 
# pwiener(dat$q[1], alpha=2, tau=.3, beta=.5, delta=.5, resp=dat$resp[1])
# 
# wiener_plot(control.data)
# 
# x <- c(2, .3, .5, .5)
# wiener_likelihood(x=x, dat=dat)
# wiener_deviance(x=x, dat=control.data)
# wiener_aic(x=x, dat=dat)
# wiener_bic(x=x, dat=dat)
# 
# # using optim, first with Nelder-Mead algorithm, then with BFGS
# optim1 <- optim(c(1, .1, .1, 1), wiener_deviance, dat=dat, method="Nelder-Mead")
# optim2 <- optim(optim1[["par"]], wiener_deviance, dat=dat, method="BFGS", hessian=TRUE)
# # using nlm, which uses a Newton-type algorithm

# # create a second data set and a list containing both data sets
# dat2 <- rwiener(n=100, alpha=2, tau=.3, beta=.5, delta=1)
# 
```